Calibrating the pixel-level Kepler imaging data with a causal data-driven model \\

### 2015-03-08 documents/cpm.tex
changes according to Hogg's comments(page1-5)

\title{
  Calibrating the pixel-level Kepler imaging data with a causal data-driven model
 \\ 
}
\author{%
  Dun~Wang\altaffilmark{\ref{CCPP}},
  Dan~Foreman-Mackey\altaffilmark{\ref{CCPP}},
  David~W.~Hogg\altaffilmark{\ref{CCPP},\ref{CDS},\ref{MPIA},\ref{email}},
  Bernhard~Sch\"olkopf\altaffilmark{\ref{MPIIS}},
  others%
}
\newcounter{address}
\setcounter{address}{1}
\altaffiltext{\theaddress}{\stepcounter{address}\label{CCPP}%
  Center for Cosmology and Particle Physics, Department of Physics, New York University}
\altaffiltext{\theaddress}{\stepcounter{address}\label{CDS}%
  Center for Data Science, New York University}
\altaffiltext{\theaddress}{\stepcounter{address}\label{MPIA}%
  Max-Planck-Institut f\"ur Astronomie, Heidelberg, Germany}
\altaffiltext{\theaddress}{\stepcounter{address}\label{email}%
  To whom correspondence should be addressed; \texttt{<david.hogg@nyu.edu>}.}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Ames}%
%%   NASA Ames Research Center}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Courant}%
%%   Courant Institute of Mathematical Sciences, New York University}
\altaffiltext{\theaddress}{\stepcounter{address}\label{MPIIS}%
  Max-Planck-Institut f\"ur Intelligente Systeme, T\"ubingen}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Oxford}%
%%   Department of Physics, Oxford University}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{CfA}%
%%   Harvard--Smithsonian Center for Astrophysics}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{UCL}%
%%   Department of Physics and Astronomy, University College London}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{CMU}%
%%   McWilliams Center for Cosmology, Carnegie Mellon University}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Caltech}%
%%   Department of Astronomy, California Institute of Technology}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Columbia}%
%%   Department of Astronomy, Columbia University}

\begin{abstract}
In general, astronomical observations are affected by several kinds of noise, each with its own causal source; 
there is photon noise, stochastic source variability, and residuals coming from imperfect calibration of the detector or telescope. 
In particular, the precision of NASA Kepler photometry for exoplanet science—
the most precise photometric measurements of stars ever made—
appears to be limited by unknown or untracked variations in spacecraft pointing and temperature, 
and unmodeled stellar variability. Here we present the Causal Pixel Model (CPM) for Kepler data, 
a data-driven model intended to capture variability but preserve transit signals. 
The CPM works at the pixel level (not the photometric measurement level); 
it can capture more fine-grained information about the variation of the spacecraft than is available in the pixel-summed aperture photometry. 
The basic idea is that CPM predicts each target pixel value from a large number of pixels of other stars sharing the instrument variabilities 
while not containing any information on possible transits at the target star. In addition, we use the target star’s future and past (auto-regression). 
By appropriately separating the data into training and test sets, we ensure that information about any transit will be perfectly isolated from the fitting of the model. 
The method has four hyper-parameters (the number of predictor stars, the auto-regressive window size, 
and two L2-regularization amplitudes for model components), which we set by cross-validation. 
We determine a generic set of hyper-parameters that works well on most of the stars and apply the method to a corresponding set of target stars. 
For these stars, we calculate the Combined Differential Photometric Precision (CDPP), which indicates the noise level of a certain time scale. 
We find that we can consistently outperform (for the purposes of exoplanet detection) the Kepler Pre-search Data Conditioning (PDC) method for exoplanet discovery.

@@

In general, there are many reasons for apparent photometric variability in a \Kepler\ source.
There is intrinsic stellar variability,
  which is of interest to some and a nuisance to others.
There is also variability of overlapping fainter stars;
  that is, confusion noise combined with variability of the confusing sources.
There are small changes in spacecraft pointing,
  which leads to slightly different illumination of the focal-plane pixels,
  and thus different sensitivity to errors (problems) in the device flat-field \Bernhard{I recall Dan saying that they don't do flat-fielding at Kepler?} or sensitivity map.
There are also \emph{intra-pixel} sensitivity variations that can contribute (\todo{cite white paper; Spitzer results}).
There are small changes in spacecraft temperature,
  which lead to point-spread function (PSF) and differential (across the focal plane) pointing changes.
These also lead to changes in pixel and intra-pixel illumination.
Stellar proper motion, geometric parallaxes, and differential stellar aberration as the spacecraft orbits all do more of the same.
There is electronic cross-talk between detectors and charge-transfer inefficiency;
  these can effectively transfer variability from one source to another.
There are additional electronics effects like ``rolling bands'' that put additional features into lightcurves.
There are also probably changes to the detector sensitivity with temperature and time,
  possibly illimunation-history effects,
  and possibly sources of variability not yet considered.
The remarkable thing about \Kepler\ is that it is trying to measure stars at a level of precision
  much higher than ever previously attempted;
  new effects really \emph{must} appear at some point.
In \figurename~\ref{fig:pixelpatch}, we show the pixel-level variability in the \Kepler\ data
  near one bright star that shows minimal intrinsic variability;
  this figure highlights the spacecraft-induced effects.

@@

The method proposed here, \name, shares the motivation of PDC.
The main differences are that
\begin{enumerate}
\item
\name\ works in the pixel domain, not the lightcurve domain, so it has access to more fine-grained information,
\item
\name has far more freedom (far more parameters) than the PDC
  but it strictly avoids over-fitting the lightcurves on exoplanet-transit time-scales
  through strong regularization and a train-and-test framework,
\item \name directly optimizes for prediction accuracy, while PDC uses PCA projections
\item
the purpose of \name\ is to search for exoplanets, so in \name, intrinsic stellar
  variability is also removed as well as the systematics.
\item \name uses as inputs only instantaneous values of other stars, plus near future and past of the star itself in an autoregressive fashion. PDC works with time series (lightcurves) of the other stars
\end{enumerate}

@@

The scope of this project is an intervention---\name---that makes the \Kepler\ transit photometry more precise.\Bernhard{doesn't this contradict our statement that we only care about transits and we remove the lightcurve variability? Suggestion: The scope of this project is an intervention---\name---that makes the \Kepler\ transit photometry more precise.}
\Dun{Yes, I think so and made this change.}
We describe the method and deliver all the relevant code in a public, open-source repository.
We also provide an interface to the \Kepler\ data that delivers improved ``\name\ photometry''
  for every \Kepler\ target.

@@

The point of these models is to be \emph{flexible},
  so the usual approach is to make the input data set very large,
  and the number of parameters large,
  often as large as---or larger than---the target data available.
Such fits require \emph{regularization} to break degeneracies
  and control ill-constrained parameters.
These regularizations do for optimizations what priors do for posterior probability inferences;
  they express the desired behavior of the fit in the absence of data
  or along directions in parameter space in which the data are not constraining.
General regularizations or priors will break the convexity of linear model fitting;
  if convexity is to be maintained, these regularizations also have to be quadratic,
  or else one of a class of other known forms (one of which is L1-regularization).\Bernhard{deleted ``small number'' -- I think it is convex for Lp with $p\ge1$.}
Given these considerations, it makes sense to try to build our pixel-level model
  using a linear model with a quadratic objective function and a quadratic regularization;
  this is what \name\ will be.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\columnwidth]{compare_10187017}
\caption{\label{hyperparameter} Comparison between default and optimized hyper-parameters. 
In the top panel, SAP flux is plotted in black, \name\ prediction of default hyper parameters is in red line, 
\name\ prediction of optimized hyper-parameters is in green line. 
The middle panel shows the CPM flux of the optimized hyper-parameters, 
while CPM flux of the default hyper-parameters is in the bottom panel, 
the two vertical blue lines indicate the location of the transit signal, and signal to noise ratio is calculated in both situation. 
The optimized hyper-parameter performs better than the default one, but the results from both situation are close.}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\columnwidth]{distortion_9822284_normal}
\caption{\label{distortion} \name\ preserves transit signals. 
Fake signal with amplitude 1.0005 is inserted into the single pixel (1, 2) of star KIC 9822284. 
In the top panel, the black points shows the SAP flux of the pixel with the injected signal, 
the green line is the fitting of the SAP flux, and the red line is the \name\ prediction. 
In the bottom panel, red and green points represent the \name\ flux and fit flux, 
while the signal value is indicated by the horizontal line. 
CPM largely preserves the signal while the fitting flux overfit the signal.}
\end{figure}

@@

\caption{\label{fluxes} Corrected fluxes using \name, for 12 example stars, spanning the main magnitude (brightness) range encountered. 
In first column plots,   we consider bright stars with magnitude from 9 to 10.5, second cloumn are stars with magnitude from 11 to 12.5,  
and the third column are faint stars with magnitude above 13. In the first two rows, we present some quiet stars. 
In all three panels, the top panel shows the SAP flux (black) and the CPM regression (red), i.e., our prediction of the star from other stars. 
The middle panel shows the CPM flux corrected using the regression, and the bottom shows the PDC flux. 
One can see that the CPM flux curve preserves the exoplanet transits (little downward spikes), 
while removing a substantial part of the variability present in the PDC flux. 
Both the bottom two rows present some variable stars with slow and fast variability. 
For the variable stars, we did not present the PDC comparison, since the PDC flux try to preserve intrinsic steller variability.}
\end{figure}

\begin{figure*}[htb]
\centering
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{pdc_cdpp12_r}
\caption{}
\end{subfigure}%
%\hfill
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{boxplot_r}
\caption{}
\end{subfigure}%
%\hfill
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{pdc_cdpp12_hist_r100}
\caption{}
\end{subfigure}
\caption{\label{cdpp} Comparison of the proposed method (CPM) to the \Kepler\ Pre-search Data Conditioning (PDC) method 
in terms of Combined Differential Photometric Precision (CDPP) (see text). 
Plot (a) shows our performance (red) vs.\ the PDC performance in a scatter plot, as a function of star magnitude 
(note that larger magnitude means fainter stars, and smaller values of CDPP indicate a higher quality as measured by CDPP.  
Plot (b) bins the same dataset and shows box plots within each bin, indicating median, top quartile and bottom quartile. 
The red box corresponds to CPM, while the black box refers to PDC. Plot (c), finally, shows a histogram of CDPP values. 
Note that the red histogram has more mass towards the left, i.e., smaller values of CDPP, indicating that our method overall outperforms PDC, the Kepler ``gold standard.'' 
}
\end{figure*}

>>>>

\title{
  Calibrating the pixel-level Kepler imaging data with a causal data-driven model
 \\ 
}
\author{%
  Dun~Wang\altaffilmark{\ref{CCPP}},
  Dan~Foreman-Mackey\altaffilmark{\ref{CCPP}},
  David~W.~Hogg\altaffilmark{\ref{CCPP},\ref{CDS},\ref{MPIA},\ref{email}},
  Bernhard~Sch\"olkopf\altaffilmark{\ref{MPIIS}}
  }
\newcounter{address}
\setcounter{address}{1}
\altaffiltext{\theaddress}{\stepcounter{address}\label{CCPP}%
  Center for Cosmology and Particle Physics, Department of Physics, New York University}
\altaffiltext{\theaddress}{\stepcounter{address}\label{CDS}%
  Center for Data Science, New York University}
\altaffiltext{\theaddress}{\stepcounter{address}\label{MPIA}%
  Max-Planck-Institut f\"ur Astronomie, Heidelberg, Germany}
\altaffiltext{\theaddress}{\stepcounter{address}\label{email}%
  To whom correspondence should be addressed; \texttt{<david.hogg@nyu.edu>}.}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Ames}%
%%   NASA Ames Research Center}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Courant}%
%%   Courant Institute of Mathematical Sciences, New York University}
\altaffiltext{\theaddress}{\stepcounter{address}\label{MPIIS}%
  Max-Planck-Institut f\"ur Intelligente Systeme, T\"ubingen}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Oxford}%
%%   Department of Physics, Oxford University}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{CfA}%
%%   Harvard--Smithsonian Center for Astrophysics}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{UCL}%
%%   Department of Physics and Astronomy, University College London}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{CMU}%
%%   McWilliams Center for Cosmology, Carnegie Mellon University}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Caltech}%
%%   Department of Astronomy, California Institute of Technology}
%% \altaffiltext{\theaddress}{\stepcounter{address}\label{Columbia}%
%%   Department of Astronomy, Columbia University}

\begin{abstract}
Astronomical observations are affected by several kinds of noise, each with its own causal source; 
there is photon noise, stochastic source variability, and residuals coming from imperfect calibration of the detector or telescope. 
The precision of NASA \Kepler\ photometry for exoplanet science ---
the most precise photometric measurements of stars ever made ---
appears to be limited by unknown or untracked variations in spacecraft pointing and temperature, 
and unmodeled stellar variability. Here we present the Causal Pixel Model (\name) for \Kepler\ data, 
a data-driven model intended to capture variability but preserve transit signals. 
The \name\ works at the pixel level; So that
it can capture more fine-grained information about the variation of the spacecraft.
The \name\ predicts each target pixel value from a large number of pixels of other stars sharing the instrument variabilities 
while not containing any information on possible transits in the target star. In addition, we use the target star’s future and past (auto-regression). 
By appropriately separating, for each data point, the data into training and test sets, we ensure that information about any transit will be perfectly isolated from the model. 
The method has four hyper-parameters (the number of predictor stars, the auto-regressive window size, 
and two L2-regularization amplitudes for model components), which we set by cross-validation. 
We determine a generic set of hyper-parameters that works well for most of the stars and apply the method to a corresponding set of target stars. 
We find that we can consistently outperform (for the purposes of exoplanet detection) the \Kepler\ Pre-search Data Conditioning (PDC) method for exoplanet discovery.

@@

In general, there are many reasons for apparent photometric variability in a \Kepler\ source.
There is intrinsic stellar variability,
  which is of interest to some and a nuisance to others.
There is also variability of overlapping fainter stars;
  that is, confusion noise combined with variability of the confusing sources.
There are small changes in spacecraft pointing,
  which leads to slightly different illumination of the focal-plane pixels,
  and thus different sensitivity to variations in the device
There are also \emph{intra-pixel} sensitivity variations that can contribute (\todo{cite white paper; Spitzer results}).
There are small changes in spacecraft temperature,
  which lead to point-spread function (PSF) and differential (across the focal plane) pointing changes.
These also lead to changes in pixel and intra-pixel illumination.
Stellar proper motion, geometric parallaxes, and differential stellar aberration as the spacecraft orbits all do more of the same.
There is electronic cross-talk between detectors and charge-transfer inefficiency;
  these can effectively transfer variability from one source to another.
There are additional electronics effects like ``rolling bands'' that put additional features into lightcurves.
There are also probably changes to the detector sensitivity with temperature and time,
  possibly illimunation-history effects,
  and possibly sources of variability not yet considered.
The remarkable thing about \Kepler\ is that it is trying to measure stars at a level of precision
  much higher than ever previously attempted;
  new effects really \emph{must} appear at some point.
In \figurename~\ref{fig:pixelpatch}, we show the pixel-level variability in the \Kepler\ data
  near one bright star that shows minimal intrinsic variability;
  this figure highlights the spacecraft-induced effects.

@@

The method proposed here, \name, shares the motivation of PDC.
The main differences are that
\begin{enumerate}
\item
\name\ works in the pixel domain, not the lightcurve domain, so it has access to more fine-grained information
\item
\name\ has far more freedom (far more parameters) than the PDC
  but it strictly avoids over-fitting the lightcurves on exoplanet-transit time-scales
  through strong regularization and a train-and-test framework.
\item 
\name\ directly optimizes for prediction accuracy, while PDC uses PCA projections.
\item
the purpose of \name\ is to search for exoplanets, so in \name, intrinsic stellar
  variability is also removed as well as the systematics.
\item 
\name\ uses as inputs only instantaneous values of other stars, plus near future and past of the star itself in an autoregressive fashion, while PDC works with time series (lightcurves) of the other stars.
\end{enumerate}

@@

The scope of this project is an intervention---\name---that makes the \Kepler\ transit photometry more precise.

We describe the method and deliver all the relevant code in a public, open-source repository.
We also provide an interface to the \Kepler\ data that delivers improved ``\name\ photometry''
  for every \Kepler\ target.

@@

The point of these models is to be \emph{flexible},
  so the usual approach is to make the input data set very large,
  and the number of parameters large,
  often as large as---or larger than---the target data available.
Such fits require \emph{regularization} to break degeneracies
  and control ill-constrained parameters.
These regularizations do for optimizations what priors do for posterior probability inferences;
  they express the desired behavior of the fit in the absence of data
  or along directions in parameter space in which the data are not constraining.
General regularizations or priors will break the convexity of linear model fitting;
  if convexity is to be maintained, these regularizations also have to be quadratic,
  or else one of a class of other known forms (one of which is L1-regularization).
Given these considerations, it makes sense to try to build our pixel-level model
  using a linear model with a quadratic objective function and a quadratic regularization;
  this is what \name\ will be.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\columnwidth]{compare_10187017}
\caption{\label{hyperparameter} Comparison between default and optimized hyper-parameters. 
In the top panel, SAP flux is plotted in black, \name\ prediction of default hyper parameters is in red line, 
\name\ prediction of optimized hyper-parameters is in green line. 
The middle panel shows the \name\ flux of the optimized hyper-parameters, 
while \name\ flux of the default hyper-parameters is in the bottom panel, 
the two vertical blue lines indicate the location of the transit signal, and signal to noise ratio is calculated in both situation. 
The optimized hyper-parameter performs better than the default one, but the results from both situation are close.}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\columnwidth]{distortion_9822284_normal}
\caption{\label{distortion} \name\ preserves transit signals. 
Fake signal with amplitude 1.0005 is inserted into the single pixel (1, 2) of star KIC 9822284. 
In the top panel, the black points shows the SAP flux of the pixel with the injected signal, 
the green line is the fitting of the SAP flux, and the red line is the \name\ prediction. 
In the bottom panel, red and green points represent the \name\ flux and fit flux, 
while the signal value is indicated by the horizontal line. 
\name\ largely preserves the signal while the fitting flux overfit the signal.}
\end{figure}

@@

\caption{\label{fluxes} Corrected fluxes using \name, for 12 example stars, spanning the main magnitude (brightness) range encountered. 
In first column plots,   we consider bright stars with magnitude from 9 to 10.5, second cloumn are stars with magnitude from 11 to 12.5,  
and the third column are faint stars with magnitude above 13. In the first two rows, we present some quiet stars. 
In all three panels, the top panel shows the SAP flux (black) and the \name\ regression (red), i.e., our prediction of the star from other stars. 
The middle panel shows the \name\ flux corrected using the regression, and the bottom shows the PDC flux. 
One can see that the \name\ flux curve preserves the exoplanet transits (little downward spikes), 
while removing a substantial part of the variability present in the PDC flux. 
Both the bottom two rows present some variable stars with slow and fast variability. 
For the variable stars, we did not present the PDC comparison, since the PDC flux try to preserve intrinsic steller variability.}
\end{figure}

\begin{figure*}[htb]
\centering
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{pdc_cdpp12_r}
\caption{}
\end{subfigure}%
%\hfill
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{boxplot_r}
\caption{}
\end{subfigure}%
%\hfill
\begin{subfigure}[htb]{0.33\columnwidth}
\includegraphics[width=\columnwidth]{pdc_cdpp12_hist_r100}
\caption{}
\end{subfigure}
\caption{\label{cdpp} Comparison of the proposed method (\name) to the \Kepler\ Pre-search Data Conditioning (PDC) method 
in terms of Combined Differential Photometric Precision (CDPP) (see text). 
Plot (a) shows our performance (red) vs.\ the PDC performance in a scatter plot, as a function of star magnitude 
(note that larger magnitude means fainter stars, and smaller values of CDPP indicate a higher quality as measured by CDPP.  
Plot (b) bins the same dataset and shows box plots within each bin, indicating median, top quartile and bottom quartile. 
The red box corresponds to \name, while the black box refers to PDC. Plot (c), finally, shows a histogram of CDPP values. 
Note that the red histogram has more mass towards the left, i.e., smaller values of CDPP, indicating that our method overall outperforms PDC, the Kepler ``gold standard.'' 
}
\end{figure*}
